---
title: "Final Project Study"
author: "Zeynep TUTAR"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy = TRUE, comment = '', out.width="80%")
```

```{r}
library(readxl) #for reading excel
library(lmtest) #for the use of DW test, autocorrelation check
library(forecast) #for linear reg for time series
library(DIMORA) #for innovation dif models (GBM, Bass Model etc.)
library(fpp2) #for datasets
library(naniar)
library(GGally)
library(ggplot2)
library(corrplot)
library(tseries)
library(rcompanion)
library(dplyr)
library(tidyr)
library(splines)
```

# EDA 

## Monthly Cinema Data

```{r}
#read the data
cinema<- read_excel("cinema_monthly.xlsx")
str(cinema)

cinema.ts <- ts(cinema[,-1], start=c(2006,1), end=c(2023,12), frequency = 12)
```

Handle Missing Values

```{r}
# Shows rows with any missing values
cinema[apply(is.na(cinema), 1, any), ]  
vis_miss(cinema)
```

```{r}
# Define the total annual values for 2013 for each column
total_2013 <- list(
  number_of_screenings = 3014642,             
  audience_admissions = 106124433,         
  audience_spending = 732299026.82
)

# Define the columns to impute
columns_to_impute <- c("number_of_screenings", "audience_admissions", 
                       "audience_spending", "avg_spending_pp")

# Create a copy of the time series object for imputation
cinema_imputed <- cinema.ts

# Filter data to include only 2011, 2012, 2014, 2015 for monthly averages
selected_years <- (time(cinema.ts) >= 2011 & time(cinema.ts) < 2013) |
                  (time(cinema.ts) >= 2014 & time(cinema.ts) < 2016)

# Loop through each column to impute missing values
for (col in columns_to_impute) {
  # Extract the column data
  column_data <- as.numeric(cinema.ts[, col])
  
  if (col == "avg_spending_pp") {
    # Calculate avg_spending_pp using the imputed audience_spending and audience_admissions
    imputed_2013 <- cinema_imputed[, "audience_spending"] / 
      cinema_imputed[, "audience_admissions"]
    # Replace the missing 2013 data with the calculated values
    column_data[time(cinema.ts) >= 2013 & time(cinema.ts) < 2014] <- 
      imputed_2013[time(cinema.ts) >= 2013 & time(cinema.ts) < 2014]
    
    # Handle missing values in 2020 and 2021
    column_data[time(cinema.ts) >= 2020 & time(cinema.ts) < 
                  2021 & is.na(column_data)] <- 6.35  # Average for 2020
    column_data[time(cinema.ts) >= 2021 & time(cinema.ts) < 
                  2022 & is.na(column_data)] <- 6.76  # Average for 2021
  } 
  else {
    # Handle cumulative features using selected years for proportions
    monthly_proportions <- colMeans(matrix(column_data[selected_years], 
                                           ncol = 12, byrow = TRUE), na.rm = TRUE)
    # Normalize to proportions
    monthly_proportions <- monthly_proportions / sum(monthly_proportions)  
    imputed_2013 <- monthly_proportions * total_2013[[col]]
    
    # Replace the missing 2013 data with the imputed values
    column_data[time(cinema.ts) >= 2013 & time(cinema.ts) < 2014] <- imputed_2013
  }
  
  # Update the imputed data in the time series
  cinema_imputed[, col] <- column_data
}

# Plot the original and imputed data for comparison
for (col in columns_to_impute) {
  par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
  
  # Plot the original data
  plot(cinema.ts[, col], type = "o", pch = 16, lty = 3, cex = 0.4,
       main = paste("Original", col), xlab = "Time", ylab = col)
  # Plot the imputed data
  plot(cinema_imputed[, col], type = "o", pch = 16, lty = 3, cex = 0.4,
       main = paste("Imputed", col), xlab = "Time", ylab = col)
  
  # Optional: Pause between plots
  #cat("Press [Enter] to see the next feature...\n")
  #readline()
}

# Reset plotting layout to default
par(mfrow = c(1, 1))

```

```{r}
plot(cinema_imputed, main="Cinema Dataset", cex.lab=0.45, type="o", pch = 16, lty = 3, cex=0.6)
```



```{r}
cinema_admissions <- cinema_imputed[,1]
cinema_screening <- cinema_imputed[,2]
cinema_spending <- cinema_imputed[,3]
avg_cinema_spending_pp <- cinema_imputed[,4]
cpi_all <- cinema_imputed[,5]
cpi_culture <- cinema_imputed[,6]
covid_cases <- cinema_imputed[,7]
```


## Annual OTT Data

```{r}
ott_annual <- read_excel("ott_cinema_annual.xlsx")
str(ott_annual)

ott_annual.ts <- ts(ott_annual[,2:3], start=2006, end=2023, frequency = 1)
```

```{r}
ott_subscriptions_annual <- ott_annual.ts[,1]
ott_spending_annual <- ott_annual.ts[,2]
```


Spline Interpolation


```{r}
# Generate a monthly date sequence
monthly_dates <- seq(as.Date("2006-01-01"), as.Date("2023-12-01"), by = "month")

# Use your extracted dataset
Year <- time(ott_subscriptions_annual)  # Extract years from the time series object
ott_subscriptions_annual <- as.numeric(ott_subscriptions_annual)  # Subscriptions data
ott_spending_annual <- as.numeric(ott_spending_annual)  # Spending data

# Perform spline interpolation for monthly data
monthly_subscriptions <- spline(
  x = Year, 
  y = ott_subscriptions_annual, 
  xout = seq(2006, 2023, length.out = length(monthly_dates))
)$y

monthly_spending <- spline(
  x = Year, 
  y = ott_spending_annual, 
  xout = seq(2006, 2023, length.out = length(monthly_dates))
)$y

# Ensure no negative values after interpolation
monthly_subscriptions[monthly_subscriptions < 0] <- 0
monthly_spending[monthly_spending < 0] <- 0

# Create a final data frame for monthly interpolated data
monthly_data <- data.frame(
  Month = monthly_dates,
  ott_subscriptions = monthly_subscriptions,
  ott_spending = monthly_spending
)

# Save the interpolated monthly dataset
write.csv(monthly_data, "interpolated_fixed_monthly_data.csv", row.names = FALSE)

# Print the first 20 rows of the final dataset for verification
print(head(monthly_data, 100))
```




```{r}
ott_subscriptions <- ts(monthly_data$ott_subscriptions, start=2006, end=2023, frequency = 12)
ott_spending <- ts(monthly_data$ott_spending, start=2006, end=2023, frequency = 12)
cinema_ott_data <- ts.union(cinema_admissions, cinema_spending, avg_cinema_spending_pp,
                            cinema_screening, ott_subscriptions, ott_spending, cpi_all,
                            cpi_culture, covid_cases)
```


```{r}
plot(cinema_ott_data, type="o", pch = 16, lty = 3, cex.lab=0.6)
plot(ott_spending, type="o", pch = 16, lty = 3, cex=0.5)
plot(ott_spending_annual, type="o", pch = 16, lty = 3, cex=0.5)
pairs(cinema_ott_data)
```



```{r}
plot(monthly_data$Month, monthly_data$ott_subscriptions, type="o", pch = 16, lty = 3, cex=0.5)
plot(monthly_data$Month, monthly_data$ott_spending, type="o", pch = 16, lty = 3, cex=0.5)
```


```{r}
#Define cor matrix
cor_matrix <- cor(as.data.frame(cinema_ott_data)[length])
# Correlation plot
corrplot(cor_matrix, method = "color", type = "upper", tl.col = "black",
         tl.srt = 45, addCoef.col = "black", number.cex = 0.5, tl.cex = 0.5)
```

```{r}
# Shows rows with any missing values
as.data.frame(cinema_ott_data)[apply(is.na(as.data.frame(cinema_ott_data)), 1, any), ]  
vis_miss(as.data.frame(cinema_ott_data))
```


## Linear Regression Model




```{r}
tsdisplay(cinema_admissions)
plot(cinema_admissions, type="o", pch = 16, lty = 3, cex=0.6)
```

Just trend and season included

```{r}
# we fit a linear model with trend and seasonality
lmts_admissions<- tslm(admissions ~ trend + season) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts_admissions), col=2, lwd=2)

summary(lmts_admissions)
anova(lmts_admissions) #provides info about analysis of variance
dwtest(lmts_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts_admissions)
```

All features included with trend and season

```{r}
# we fit a linear model with trend and seasonality
lmts2_admissions<- tslm(admissions ~ trend + season + screening + spending + 
                          avg_spending_pp + cpi_all + cpi_culture + covid_cases_monthly) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts2_admissions), col=2, lwd=2)

summary(lmts2_admissions)
anova(lmts2_admissions) #provides info about analysis of variance
dwtest(lmts2_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts2_admissions)

```


Piecewise feature selection (with trend and season)

```{r}
# we fit a linear model with trend and seasonality
lmts3_admissions<- tslm(admissions ~ trend + season + screening + spending + 
                          avg_spending_pp + cpi_all + covid_cases_monthly) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts3_admissions), col=2, lwd=2)

summary(lmts3_admissions)
anova(lmts3_admissions) #provides info about analysis of variance
dwtest(lmts3_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts3_admissions)

```


```{r}
# we fit a linear model with trend and seasonality
lmts4_admissions<- tslm(admissions ~ trend + season + screening + spending + 
                          avg_spending_pp + cpi_all) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts4_admissions), col=2, lwd=2)

summary(lmts4_admissions)
anova(lmts4_admissions) #provides info about analysis of variance
dwtest(lmts4_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts4_admissions)

```



```{r}
# we fit a linear model with trend and seasonality
lmts5_admissions<- tslm(admissions ~ trend + season + spending + 
                          avg_spending_pp + cpi_all) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts5_admissions), col=2, lwd=2)

summary(lmts5_admissions)
anova(lmts5_admissions) #provides info about analysis of variance
dwtest(lmts5_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts5_admissions)

```



```{r}
# we fit a linear model with trend and seasonality
lmts6_admissions<- tslm(admissions ~ trend + season + spending + 
                          avg_spending_pp) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts6_admissions), col=2, lwd=2)

summary(lmts6_admissions)
anova(lmts6_admissions) #provides info about analysis of variance
dwtest(lmts6_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts6_admissions)

```



```{r}
# we fit a linear model with trend and seasonality
lmts7_admissions<- tslm(admissions ~ trend + season + spending + cpi_all) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts7_admissions), col=2, lwd=2)

summary(lmts7_admissions)
anova(lmts7_admissions) #provides info about analysis of variance
dwtest(lmts7_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts7_admissions)

```



```{r}
# we fit a linear model with trend and seasonality
lmts8_admissions<- tslm(admissions ~ trend + season + screening + spending 
                        + cpi_all) 

#plot of the model
plot(admissions, main="tslm_admissions", type="o", pch = 16, cex=0.4)
lines(fitted(lmts8_admissions), col=2, lwd=2)

summary(lmts8_admissions)
anova(lmts8_admissions) #provides info about analysis of variance
dwtest(lmts8_admissions) #indicates a problem of positive autocorrelation

#check the residuals
checkresiduals(lmts8_admissions)

```



```{r}
compareLM(lmts_admissions, lmts2_admissions, lmts3_admissions, lmts4_admissions,
          lmts5_admissions, lmts6_admissions, lmts7_admissions, lmts8_admissions)
```



```{r}
pairwiseModelAnova(lmts_admissions, lmts2_admissions, lmts3_admissions, 
                   lmts4_admissions, lmts5_admissions, lmts6_admissions, 
                   lmts7_admissions, lmts8_admissions)
```


Forecasting
: this part doesnt work because i need future values of other predictor (screening, spending etc.). For that, I can use ARIMA.
Also the residuals are not WN so still forecast anyway? but i did forecasting with diff models anyway so why not? TODO

```{r}
forcast_admissions <- forecast(lmts_admissions, h=24)
forcast_admissions
plot(forcast_admissions)
lines(fitted(lmts4_admissions), col=2, lwd=2)
```

## Bass Model (BM)


```{r}
# Fit the Bass Model
BM.cinema_admissions <- BM(cinema_admissions)
summary(BM.cinema_admissions)

checkresiduals(BM.cinema_admissions)
plot.Dimora(BM.cinema_admissions)
```
Future Prediction

```{r}
# Generate time index for observed data (2006 to 2023, monthly)
observed_months <- seq(as.Date("2006-01-01"), as.Date("2023-12-01"), by = "month")
# Generate time index for predictions (2006 to 2033, monthly)
predicted_months <- seq(as.Date("2006-01-01"), as.Date("2033-12-01"), by = "month")
```


```{r}
# Predict 10 years into the future
pred_BM.cinema_admissions<- predict(BM.cinema_admissions, newx = 1:336) 
inst_pred_BM.cinema_admissions<- make.instantaneous(pred_BM.cinema_admissions)

par(mfrow = c(1, 2))
# Plot observed cumulative cinema_admissions
plot(observed_months, cumsum(cinema_admissions), type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Cumulative Cinema Admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cumsum(cinema_admissions), pred_BM.cinema_admissions)))
# Overlay predicted cumulative cinema_admissions
lines(predicted_months, pred_BM.cinema_admissions, lwd = 2, col = 2)
legend("bottomright", c("Observed", "Predicted"), col = c(1, 2),
       lty = c(3, 1), lwd = c(1, 2), pch = c(16, NA), merge = TRUE, cex=0.8)


# Plot observed monthly cinema_admissions
plot(observed_months, cinema_admissions, type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Monthly Cinema Admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cinema_admissions, inst_pred_BM.cinema_admissions)))
# Overlay predicted instantaneous cinema_admissions
lines(predicted_months, inst_pred_BM.cinema_admissions, lwd = 3, col = 2)
legend("topright", c("Observed", "Predicted"), col = c(1, 2),
       lty = c(3, 1), lwd = c(1, 3), pch = c(16, NA), merge = TRUE, cex=0.8)

par(mfrow = c(1, 1))
```


## GBM

```{r}
# GBM exponential shock
GBM.cinema_admissions<- GBM(cinema_admissions,shock = "exp",nshock = 1,
                                 prelimestimates = c(1.963349e+09, 4.244444e-03, 
                                                     9.822524e-03, 170,-0.1,-1)) 
# m,p,q, a(starting point), b(memory of the shock), c(intensity) 
summary(GBM.cinema_admissions)

checkresiduals(GBM.cinema_admissions)
plot.Dimora(GBM.cinema_admissions)
```

```{r}
res_GBM <- residuals(GBM.cinema_admissions)
dwtest(res_GBM)
```


Future Prediction

```{r}
# Predict 10 years into the future
pred_GBM.cinema_admissions<- predict(GBM.cinema_admissions, newx = 1:336) 
inst_pred_GBM.cinema_admissions<- make.instantaneous(pred_GBM.cinema_admissions)

par(mfrow = c(1, 2))
# Plot observed cumulative cinema_admissions
plot(observed_months, cumsum(cinema_admissions), type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Cumulative cinema_admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cumsum(cinema_admissions), pred_GBM.cinema_admissions)))
# Overlay predicted cumulative cinema_admissions
lines(predicted_months, pred_GBM.cinema_admissions, lwd = 2, col = 3)
legend("bottomright", c("Observed", "Predicted"), col = c(1, 3),
       lty = c(3, 1), lwd = c(1, 2), pch = c(16, NA), merge = TRUE, cex=0.8)


# Plot observed monthly cinema_admissions
plot(observed_months, cinema_admissions, type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Monthly cinema_admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cinema_admissions, inst_pred_GBM.cinema_admissions)))
# Overlay predicted instantaneous cinema_admissions
lines(predicted_months, inst_pred_GBM.cinema_admissions, lwd = 3, col = 3)
legend("topright", c("Observed", "Predicted"), col = c(1, 3),
       lty = c(3, 1), lwd = c(1, 3), pch = c(16, NA), merge = TRUE, cex=0.8)

par(mfrow = c(1, 1))
```
## GGM

```{r}
GGM.cinema_admissions<- GGM(cinema_admissions, prelimestimates=c(3.496741e+09, 0.001, 0.01, 
                                                  2.693658e-03, 2.973322e-03))
summary(GGM.cinema_admissions)

checkresiduals(GGM.cinema_admissions)
plot.Dimora(GGM.cinema_admissions)
```

Future Prediction

```{r}
# Predict 10 years into the future
pred_GGM.cinema_admissions<- predict(GGM.cinema_admissions, newx = 1:336) 
inst_pred_GGM.cinema_admissions<- make.instantaneous(pred_GGM.cinema_admissions)

par(mfrow = c(1, 2))
# Plot observed cumulative cinema_admissions
plot(observed_months, cumsum(cinema_admissions), type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Cumulative cinema_admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cumsum(cinema_admissions), pred_GGM.cinema_admissions)))
# Overlay predicted cumulative cinema_admissions
lines(predicted_months, pred_GGM.cinema_admissions, lwd = 2, col = "deepskyblue1")
legend("bottomright", c("Observed", "Predicted"), col = c(1, "deepskyblue1"),
       lty = c(3, 1), lwd = c(1, 2), pch = c(16, NA), merge = TRUE, cex=0.8)


# Plot observed monthly cinema_admissions
plot(observed_months, cinema_admissions, type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Monthly cinema_admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cinema_admissions, inst_pred_GGM.cinema_admissions)))
# Overlay predicted instantaneous cinema_admissions
lines(predicted_months, inst_pred_GGM.cinema_admissions, lwd = 3, col = "deepskyblue1")
legend("topright", c("Observed", "Predicted"), col = c(1, "deepskyblue1"),
       lty = c(3, 1), lwd = c(1, 3), pch = c(16, NA), merge = TRUE, cex=0.8)

par(mfrow = c(1, 1))
```

```{r}
# Plot observed monthly cinema_admissions
plot(observed_months, cinema_admissions, type = "o", pch = 16, lty = 3, 
     cex = 0.6, xlab = "Time", ylab = "Monthly Cinema Admissions",
     xlim = range(c(observed_months, predicted_months)), 
     ylim = c(0, max(cinema_admissions, inst_pred_GGM.cinema_admissions)))

# Overlay predicted instantaneous cinema_admissions
lines(predicted_months, inst_pred_BM.cinema_admissions, lwd = 3, col = 2) 
lines(predicted_months, inst_pred_GBM.cinema_admissions, lwd = 3, col = 3)
lines(predicted_months, inst_pred_GGM.cinema_admissions, lwd = 3, col = "deepskyblue1")
legend("topright", c("BM", "GBM", "GGM"), col = c(2, 3, "deepskyblue1"),
       lty = c(1, 1, 1), lwd = c(3, 3, 3), merge = TRUE, cex=0.8)
```
## ARIMA

```{r}
tsdisplay(cinema_admissions)
```

General indication (for differenced data): 
- If the ACF is exponentially decaying or sinusoidal and there is a significant spike at lag p in PACF and nothing else, it may be an ARIMA(p,d,0). 
- If the PACF is exponentially decaying or sinusoidal and there is a significant spike at lag p in ACF and nothing else, it may be an ARIMA(0,d,q). 


```{r}
diff1<- diff(cinema_admissions)
tsdisplay(diff1, main = "cinema_admissions (d=1)")
```

```{r}
diff12<- diff(diff1, lag=12) 
tsdisplay(diff12, main = "cinema_admissions (d=1, D=1, s=12)")
```


```{r}
# first Arima model 
arima1.cinema_admissions<- Arima(cinema_admissions, order=c(2,1,2), seasonal=c(1,1,1))
summary(arima1.cinema_admissions)
fit.arima1.cinema_admissions<- fitted(arima1.cinema_admissions)

plot(cinema_admissions, type="o", pch = 16, cex=0.6)
lines(fit.arima1.cinema_admissions, col="green", lwd=2)


forecast.arima1.cinema_admissions<- forecast(arima1.cinema_admissions)
plot(forecast.arima1.cinema_admissions)

res.arima1.cinema_admissions<- residuals(arima1.cinema_admissions)
tsdisplay(res.arima1.cinema_admissions)

checkresiduals(arima1.cinema_admissions)
```

```{r}
# second Arima model 
arima2.cinema_admissions<- Arima(cinema_admissions, order=c(9,1,2), seasonal=c(1,1,1))
summary(arima2.cinema_admissions)
fit.arima2.cinema_admissions<- fitted(arima2.cinema_admissions)

plot(cinema_admissions, type="o", pch = 16, cex=0.6)
lines(fit.arima2.cinema_admissions, col="blue", lwd=2)

forecast.arima2.cinema_admissions<- forecast(arima2.cinema_admissions)
plot(forecast.arima2.cinema_admissions)

res.arima2.cinema_admissions<- residuals(arima2.cinema_admissions)
tsdisplay(res.arima2.cinema_admissions)

checkresiduals(arima2.cinema_admissions)

```

```{r}
# Auto ARIMA
auto.arima.cinema_admissions<- auto.arima(cinema_admissions)
summary(auto.arima.cinema_admissions)
fit.auto.arima.cinema_admissions<- fitted(auto.arima.cinema_admissions)

plot(cinema_admissions, type="o", pch = 16, cex=0.6)
lines(fit.auto.arima.cinema_admissions, col="red", lwd=2)

forecast.auto.arima.cinema_admissions<- forecast(auto.arima.cinema_admissions)
autoplot(forecast.auto.arima.cinema_admissions)

res.auto.arima.cinema_admissions<- residuals(auto.arima.cinema_admissions)
tsdisplay(res.auto.arima.cinema_admissions)

checkresiduals(auto.arima.cinema_admissions)
```

```{r}
AIC(arima1.cinema_admissions, arima2.cinema_admissions, auto.arima.cinema_admissions)
```

TODO: mix modeller yap, tek baĹźÄ±na arima yeterisz, covid patlamasÄ±nÄ± gĂ¶remiyor

## Exponential Smoothing

### 1) Simple exponential smoothing

```{r}
SES1.cinema_admissions<- ses(cinema_admissions, alpha=0.2, initial="simple", h=120)
# alpha: # Value of smoothing parameter for the level. 
# h: Number of periods for forecasting.

# Replace negative lower bounds with zero
SES1.cinema_admissions$lower[SES1.cinema_admissions$lower < 0] <- 0
summary(SES1.cinema_admissions)
accuracy(SES1.cinema_admissions)
```


```{r}
SES2.cinema_admissions<- ses(cinema_admissions, alpha=0.6, initial="simple", h=120)
# Replace negative lower bounds with zero
SES2.cinema_admissions$lower[SES1.cinema_admissions$lower < 0] <- 0
summary(SES2.cinema_admissions)
```



```{r}
SES3.cinema_admissions<- ses(cinema_admissions, alpha=0.9, h=120)
# Replace negative lower bounds with zero
SES3.cinema_admissions$lower[SES3.cinema_admissions$lower < 0] <- 0
summary(SES3.cinema_admissions)
```


```{r}
par(mfrow=c(3,1), pin=c(5,1))
plot(cinema_admissions, ylab="cinema_admissions\n(alpha=0.2)", type="o", pch = 16, cex=0.6)
lines(fitted(SES1.cinema_admissions), col="green", lwd=2)

plot(cinema_admissions, ylab="cinema_admissions\n(alpha=0.6)", type="o", pch = 16, cex=0.6)
lines(fitted(SES2.cinema_admissions), col="blue", lwd=2)

plot(cinema_admissions, ylab="cinema_admissions\n(alpha=0.9)", type="o", pch = 16, cex=0.6)
lines(fitted(SES3.cinema_admissions), col="red", lwd=2)

par(mfrow=c(1,1))

```

```{r}
SES1.cinema_admissions$model$aic
SES2.cinema_admissions$model$aic
SES3.cinema_admissions$model$aic
```


```{r}
autoplot(cinema_admissions)+
  autolayer(SES3.cinema_admissions, series="alpha=0.9", PI=T)+
  autolayer(SES1.cinema_admissions, series="alpha=0.2", PI=F)+
  autolayer(SES2.cinema_admissions, series="alpha=0.6", PI=F)+
  ylab("Cinema Admissions")+xlab("Year")
```

### 2) Holt's method (for Trend)

```{r}

holt1.cinema_admissions<- holt(cinema_admissions, h=120)
holt2.cinema_admissions<- holt(cinema_admissions, damped=T, h=120)
# Replace negative lower bounds with zero
holt1.cinema_admissions$lower[holt1.cinema_admissions$lower < 0] <- 0
holt2.cinema_admissions$lower[holt2.cinema_admissions$lower < 0] <- 0

summary(holt1.cinema_admissions)
summary(holt2.cinema_admissions)

autoplot(cinema_admissions)+
  autolayer(holt1.cinema_admissions, series="Holt's method", PI=F)+
  autolayer(holt2.cinema_admissions, series="Damped Holt's method", PI=F)

```

### 3) Holt-Winters method (for Trend and Seasonality)

```{r}
hw.cinema_admissions<- hw(cinema_admissions, seasonal="additive")
# Replace negative lower bounds with zero
hw.cinema_admissions$lower[hw.cinema_admissions$lower < 0] <- 0

summary(hw.cinema_admissions)

autoplot(hw.cinema_admissions)+
  autolayer(fitted(hw.cinema_admissions), series="Holt-Winters' additive")+
  ylab("Cinema Admissions")+xlab("Year")
```

```{r}
SES3.cinema_admissions$model$aicc
holt1.cinema_admissions$model$aicc
holt2.cinema_admissions$model$aicc
hw.cinema_admissions$model$aicc
```


# Annual OTT and Cinema Data


## EDA 


```{r}
plot(ott_cinema_annual.ts, type="o", pch = 16, lty = 3, cex.lab=0.5)
```

Handle Missing Values

```{r}
# Shows rows with any missing values
ott_cinema_annual[apply(is.na(ott_cinema_annual), 1, any), ]  
```

```{r}
SES.households_internet <- holt(households_internet, damped=T, h=1)

summary(SES.households_internet)

autoplot(SES.households_internet)+
  autolayer(fitted(SES.households_internet), series="Damped Holt's method")+
  ylab("Households with Internet (%)")+xlab("Year")
```

```{r}
households_internet <- c(households_internet[-length(households_internet)], 
                         SES.households_internet$mean)

ott_cinema_annual.ts[, "households_w_internet"] <- households_internet

ott_cinema_annual$households_w_internet[is.na(ott_cinema_annual$households_w_internet)] <- SES.households_internet$mean
```



```{r}
col = colnames(ott_cinema_annual.ts)
ott_cinema_annual[,col[-1]] %>%
  GGally::ggpairs()
```



```{r}
#Define cor matrix
ott_cinema_annual_cormatrix <- cor(ott_cinema_annual[,-1])
# Correlation plot
corrplot(ott_cinema_annual_cormatrix, method = "color", type = "upper", tl.col = "black",
         tl.srt = 45, addCoef.col = "black", number.cex = 0.5, tl.cex = 0.5)
```

## Diffiusion Models for OTT Subscriptions and Cinema Admissions

Perform a competition study between OTT and Cinema

```{r}
# Apply Min-Max Normalization
normalize_minmax <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}
ott_subscriptions.norm <- normalize_minmax(ott_subscriptions)
cinema_admissions.norm <- normalize_minmax(cinema_admissions)
```


```{r}
# Plot the normalized time series
plot(cinema_admissions.norm, type = "o", pch = 16, lty = 3, col = "blue", 
     xlab = "Year", ylab = "Normalized Values",
     main = "Normalized OTT Subscriptions and Cinema Admissions", ylim = c(0, 1))
points(ott_subscriptions.norm, type = "o", pch = 16, lty = 3, col = "red")
legend("left", legend = c("Cinema Admissions", "OTT Subscriptions"), 
       col = c("blue", "red"), lty = 3, pch = 16, cex=0.7)
```

Estimating initial parameters with Bass Model:

```{r}
# Cinema Admissions
BM.cinema_admissions<-BM(cinema_admissions, display = T)
m.c = BM.cinema_admissions$Estimate[1, 1]     # Market potential for Cinema
p.c = BM.cinema_admissions$Estimate[2, 1]    # Innovation coefficient for Cinema
q.c = BM.cinema_admissions$Estimate[3, 1]    # Imitation coefficient for Cinema
summary(BM.cinema_admissions)
checkresiduals(BM.cinema_admissions)

# OTT Subscriptions
BM.ott_subscriptions<-BM(ott_subscriptions, display = T)
m.o = BM.ott_subscriptions$Estimate[1, 1]     # Market potential for OTT
p.o = BM.ott_subscriptions$Estimate[2, 1] # Innovation coefficient for OTT
q.o = BM.ott_subscriptions$Estimate[3, 1]  # Imitation coefficient for OTT
summary(BM.ott_subscriptions)
checkresiduals(BM.ott_subscriptions)
```

GBM for cinema admissions:
```{r}
# Cinema Admissions
# GBM exponential shock

# Define the time period when the COVID-19 shock begins
covid_start <- 2020 - 2006 + 1

# m,p,q, a(starting point), b(memory of the shock), c(intensity)
GBM.cinema_admissions<- GBM(cinema_admissions, shock = "exp", nshock = 1,
                     prelimestimates = c(m.c, p.c, q.c, covid_start, -0.5, -0.1)) 

summary(GBM.cinema_admissions)
checkresiduals(GBM.cinema_admissions)

m.c = BM.cinema_admissions$Estimate[1, 1]    # Market potential for Cinema
p.c = BM.cinema_admissions$Estimate[2, 1]    # Innovation coefficient for Cinema
q.c = BM.cinema_admissions$Estimate[3, 1]    # Imitation coefficient for Cinema
```
GGM for OTT subscriptions:

```{r}
GGM.ott_subscriptions<- GGM(ott_subscriptions, 
                            prelimestimates=c(m.o, 0.01, 0.01, p.o, q.o), oos=10)
summary(GGM.ott_subscriptions)
checkresiduals(GGM.ott_subscriptions)
```



```{r}
###estimate the UCRCD (with delta and gamma)
UCRCD.adm.subs<- UCRCD(
  series1 = cinema_admissions,  # Instantaneous cinema admissions
  series2 = ott_subscriptions,  # Instantaneous OTT subscriptions
  m1 = m.c,                    # Market potential for cinema (GBM result)
  p1c = p.c,                   # Innovation coefficient for cinema
  q1c = q.c,                   # Imitation coefficient for cinema
  m2 = m.o,                    # Market potential for OTT (BM result)
  p2 = p.o,                   # Innovation coefficient for OTT
  q2 = q.o                    # Imitation coefficient for OTT
)
summary(UCRCD.adm.subs)
```


```{r}
#Expenses
plot(cinema_audience_spending,xlab="year", ylab="Expenses", type= "o", 
     pch=16, lty=3, cex=0.7, ylim=c(0, max(cinema_audience_spending, ott_spending)))
points(ott_spending, type= "o", pch=16, lty=3, cex=0.7, col=2)
```
#TODO: scaling ile subs ve adm karĹźÄ±laĹźtÄ±rabilirsin. https://rdrr.io/github/federicozanghi/DIMORA/man/UCRCD.html bu linkteki methodla GBM shockunu verebilirsin. bunun iĂ§in de Ă¶nde datalarÄ± BM, GBM ve GGM ile dene.

```{r}
###estimate the UCRCD (with delta and gamma)
UCRCD.oc.exp<- UCRCD(cinema_audience_spending, ott_spending, display=T)
summary(UCRCD.oc.exp)
```


```{r}
##we make a plot of the UCRCD model 
plot(cinema_audience_spending,xlab="year", ylab="Expenses", type= "o", 
     pch=16, lty=3, cex=0.7, ylim=c(0, max(cinema_audience_spending, ott_spending)))
points(ott_spending, type= "o", pch=16, lty=3, cex=0.7, col=2)

observed_years <- seq(2006, 2023, by = 1)
lines(observed_years, (UCRCD.oc.exp$fitted.i[[1]]), lwd=2, col=1)
lines(observed_years, (UCRCD.oc.exp$fitted.i[[2]]), lwd=2, col=2)

###we also add a line and some text within the plot
abline(v=2020, lty=2)
text(2020, 0.8, pos=3, "Covid-19 Pandemic")
```






```{r}
#we estimate a simple Bass Model 
BM.ott_subscriptions <- BM(ott_subscriptions, display = T)
summary(BM.ott_subscriptions)

#we are not sure if this is the maximum peak since we are still in the increasing 
#part of the process. So, simple Bass Model is probably not enough. But still, 
#start with Bass Model as a starting point and then move on to more complex models
#Looks like Bass Model cannot capture the data well. 
#while there is still an increase in the data, the model predicts a decrease.
```


```{r}
Acf(ott_subscriptions)
Acf(ott_spending)
Acf(cinema_admissions)
Acf(cinema_audience_spending)
```


```{r}
# Predict 80 years into the future
pred.ott_subs <- predict(BM.ott_subs, newx = 1:98) # Extend 80 years beyond 18 years of data
instant.ott_subs <- make.instantaneous(pred.ott_subs)

# Generate time index for observed data (2006 to 2023)
observed_years <- seq(2006, 2023, by = 1)
# Generate time index for predictions (2006 to 2103)
pred_years <- seq(2006, 2006 + 97, by = 1)

# Plot observed cumulative subscriptions
plot(observed_years, cumsum(ott_subs), type = "o", xlab = "Year", ylab = "Cumulative subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2103), ylim = c(0, 1000))
# Overlay the predicted cumulative subscriptions
lines(pred_years, pred.ott_subs, lwd = 2, col = 2)

# Plot annual subscriptions
plot(observed_years, ott_subs, type = "o", xlab = "Year", ylab = "Annual Subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2103), ylim = c(0, max(ott_subs, instant.ott_subs)))

# Overlay the predicted instantaneous subscriptions
lines(pred_years, instant.ott_subs, lwd = 2, col = 2)


res.fit.ott_subs<- residuals(BM.ott_subs)
plot(res.fit.ott_subs, type = "o", xlab = "Year", ylab = "Annual Subs",
     pch = 16, lty = 3, cex = 0.6)
Acf(res.fit.ott_subs)

res.fit.ott_exp<- residuals(BM.ott_exp)
plot(res.fit.ott_exp, type = "o", xlab = "Year", ylab = "Annual Subs",
     pch = 16, lty = 3, cex = 0.6)
Acf(res.fit.ott_exp)

#----------------------------------------------------------------

BM.cinema_adm<-BM(cinema_adm,display = T)
summary(BM.cinema_adm)
BM.cinema_exp<-BM(cinema_exp,display = T)
summary(BM.cinema_exp)


#------------------------------------------------------
# OTT GBM

# Fit the Bass Model
BM.ott_exp <- BM(ott_exp)
summary(BM.ott_exp) 
res.BM.ott_exp<- residuals(BM.ott_exp)
plot(res.BM.ott_exp, type = "o", pch = 16, lty = 3, cex = 0.6)
Acf(res.BM.ott_exp)

# GBM exponential shock
GBM.ott_exp<- GBM(ott_exp,shock = "rett",nshock = 1,
                     prelimestimates = c(8.894847e+04, 4.134324e-05, 5.540857e-01, 15, 17, -1)) 
# m,p,q, a(starting point), b(memory of the shock), c(intensity)
summary(GBM.ott_exp)

res.GBM.ott_exp<- residuals(GBM.ott_exp)
plot(res.GBM.ott_exp, type = "o", pch = 16, lty = 3, cex = 0.6)
Acf(res.GBM.ott_exp)

#when the peak is sooo far away, we have a situation that we havent reached the 
#peak point yet. So the market potential is not reliable at this point. So we 
#can rely on the model for a short period of estimations.

# Predict 50 years into the future
pred_GBM.ott_exp<- predict(GBM.ott_exp, newx = 1:48) # Extend 50 years beyond 18 years of data
pred_GBM.ott_exp_inst<- make.instantaneous(pred_GBM.ott_exp)

# Generate time index for observed data 
observed_years <- seq(2006, 2023, by = 1)
# Generate time index for predictions 
pred_years <- seq(2006, 2006 + 47, by = 1)

# Plot observed cumulative subscriptions
plot(observed_years, cumsum(ott_exp), type = "o", xlab = "Year", ylab = "Cumulative subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 47), ylim = c(0, max(cumsum(ott_exp), pred_GBM.ott_exp)))
# Overlay the predicted cumulative subscriptions
lines(pred_years, pred_GBM.ott_exp, lwd = 2, col = 2)

# Plot annual subscriptions
plot(observed_years, ott_exp, type = "o", xlab = "Year", ylab = "Annual Subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 47), ylim = c(0, max(ott_exp, pred_GBM.ott_exp_inst)))

# Overlay the predicted instantaneous subscriptions
lines(pred_years, pred_GBM.ott_exp_inst, lwd = 2, col = 2)

#------------------------------------------------------
# Cinema GBM

# Fit the Bass Model
BM.cinema_adm <- BM(cinema_adm)
summary(BM.cinema_adm) 

# GBM exponential shock
GBM.cinema_adm<- GBM(cinema_adm, shock = "exp",nshock = 1,
                     prelimestimates = c(1.979387e+03, 5.055723e-02, 1.157916e-01, 15,-0.1,-1)) 
# m,p,q, a(starting point), b(memory of the shock), c(intensity)
summary(GBM.cinema_adm)

res.GBM.cinema_adm <- residuals(GBM.cinema_adm)
plot(res.GBM.cinema_adm, type = "o", pch = 16, lty = 3, cex = 0.6)
Acf(res.GBM.cinema_adm)

# Predict 50 years into the future
pred_GBM.cinema_adm<- predict(GBM.cinema_adm, newx = 1:68) # Extend 50 years beyond 18 years of data
inst_pred_GBM.cinema_adm<- make.instantaneous(pred_GBM.cinema_adm)

# Generate time index for observed data 
observed_years <- seq(2006, 2023, by = 1)
# Generate time index for predictions
pred_years <- seq(2006, 2006 + 67, by = 1)

# Plot observed cumulative subscriptions
plot(observed_years, cumsum(cinema_adm), type = "o", xlab = "Year", ylab = "Cumulative adm",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 67), ylim = c(0, max(cumsum(cinema_adm), pred_GBM.cinema_adm)))
# Overlay the predicted cumulative subscriptions
lines(pred_years, pred_GBM.cinema_adm, lwd = 2, col = 2)

# Plot annual subscriptions
plot(observed_years, cinema_adm, type = "o", xlab = "Year", ylab = "Annual adm",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 67), ylim = c(0, max(cinema_adm, inst_pred_GBM.cinema_adm)))

# Overlay the predicted instantaneous subscriptions
lines(pred_years, inst_pred_GBM.cinema_adm, lwd = 2, col = 2)




#-----------------------------------------------------------------------------
#Cinema(annually) GGM

BM.cinema_adm <- BM(cinema_adm)
summary(BM.cinema_adm) 
GBM.cinema_adm<- GBM(cinema_adm, shock = "exp",nshock = 1,
                     prelimestimates = c(1.979387e+03, 5.055723e-02, 1.157916e-01, 15,-0.1,-1)) 
summary(GBM.cinema_adm)

GGM.cinema_adm<- GGM(cinema_adm, prelimestimates=c(3769.43430149, 0.001, 0.01, 0.03001907, 0.03134832))
summary(GGM.cinema_adm)

res.GGM.cinema_adm <- residuals(GGM.cinema_adm)
plot(res.GGM.cinema_adm, type = "o", pch = 16, lty = 3, cex = 0.6)
Acf(res.GGM.cinema_adm)

#------------------------------------------------------
# OTT GGM

BM.ott_exp <- BM(ott_exp)
summary(BM.ott_exp) 
GBM.ott_exp<- GBM(ott_exp,shock = "rett",nshock = 1,
      prelimestimates = c(8.894847e+04, 4.134324e-05, 5.540857e-01, 15, 17, -1)) 
summary(GBM.ott_exp)

GGM.ott_exp<- GGM(ott_exp, prelimestimates=c(1.059095e+04, 0.001, 0.01, 3.870570e-05, 5.421316e-01))
summary(GGM.ott_exp)

res.GGM.ott_exp<- residuals(GBM.ott_exp)
plot(res.GGM.ott_exp, type = "o", pch = 16, lty = 3, cex = 0.6)
Acf(res.GGM.ott_exp)

# Predict 50 years into the future
pred_GGM.ott_exp<- predict(GGM.ott_exp, newx = 1:48) # Extend 50 years beyond 18 years of data
pred_GGM.ott_exp_inst<- make.instantaneous(pred_GGM.ott_exp)

# Generate time index for observed data 
observed_years <- seq(2006, 2023, by = 1)
# Generate time index for predictions 
pred_years <- seq(2006, 2006 + 47, by = 1)

# Plot observed cumulative subscriptions
plot(observed_years, cumsum(ott_exp), type = "o", xlab = "Year", ylab = "Cumulative subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 47), ylim = c(0, max(cumsum(ott_exp), pred_GGM.ott_exp)))
# Overlay the predicted cumulative subscriptions
lines(pred_years, pred_GGM.ott_exp, lwd = 2, col = 2)

# Plot annual subscriptions
plot(observed_years, ott_exp, type = "o", xlab = "Year", ylab = "Annual Subs",
     pch = 16, lty = 3, cex = 0.6, xlim = c(2006, 2006 + 47), ylim = c(0, max(ott_exp, pred_GGM.ott_exp_inst)))

# Overlay the predicted instantaneous subscriptions
lines(pred_years, pred_GGM.ott_exp_inst, lwd = 2, col = 2)
lines(pred_years, pred_GBM.ott_exp_inst, lwd = 2, col = 3)

#------------------------------------------------------


#-----------------------------------------------------------------------------

```

