---
title: "Business, Economic and Financial Data Final Project" 
subtitle: "A Time Series Analysis of Cinema Admissions in Italy"
author: "Zeynep TUTAR - 2106038"
bibliography: "Bibliography.bib"
toc: true
toc-own-page: true
df_print: kable
fontsize: 10pt
lang: en-GB
titlepage: true
titlepage-background: "img/unipdbg.pdf"
header-center: "\\leftmark"
header-right: "BEFD Final Project"
footer-left: "\\hspace{1cm}"
footer-center: "Page \\thepage"
footer-right: "\\hspace{1cm}"
natbib: true
hyperrefoptions:
  - linktoc=all
#biblio-style: apsr
colorlinks: true

editor_options: 
  markdown: 
    wrap: 72
    

header-includes:
   - \usepackage{setspace}
   - \newpage
   - \usepackage{hyperref}
   - \usepackage{multirow}
   - \usepackage{amsmath}
   - \usepackage{algorithm}
   - \usepackage{algorithmic}
   - \usepackage{authblk}
   - \usepackage{graphicx}
   - 
#   - \usepackage{fancyvrb}
#   - \RecustomVerbatimEnvironment{Verbatim}{BVerbatim}{breaklines=true,breakanywhere=true}



output:
  pdf_document:
    template: "template/eisvogel.tex"
    citation_package: natbib
    number_sections: true
    keep_tex: false
    fig_caption: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, message = FALSE,
                      results = "hide",
                      tidy.opts = list(width.cutoff = 72),
                      tidy = TRUE, comment = '',
                      fig.path='figs/', dev = "png",
                      cache = TRUE, cache.path = 'Project_Report_cache/',
                      out.width="50%")
# fig.keep = 'all'
```

```{r libraries}
library(car)
library(dplyr)
library(tidyr)
library(prophet)
library(knitr)
```


# Introduction

Over the past few years, we have grown accustomed to watching movies on
OTT (Over-The-Top) streaming—services that deliver content directly to
viewers over the internet. These platforms have become an integral part
of our daily routines, allowing us to enjoy movies in the comfort of our
homes. However, this shift has made us forget the unique joy of sharing
emotions with a crowd in a movie theater.

Particularly after the COVID-19 pandemic, it seems like the streaming
platforms became dominant, and the overwhelming volume of content on
these services has made it increasingly difficult for individual films
to draw attention in cinemas. This led me to wonder: is traditional
cinema culture approaching its end?

To explore further, this study focuses on the time series analysis and
forecasting of cinema admissions in Italy.

## Problem Definition

This study aims to address the following questions:

\begin{itemize}
    \item Is traditional cinema culture coming to an end?
    \item Will OTT services dominate the distribution of movies in the future?
    \item Have the COVID-19 pandemic and/or OTT platforms impacted theater admissions significantly?
\end{itemize}

## The Data Sources

The dataset used in this study consists of monthly data spanning from
January 2006 to December 2023, complemented by a simplified annual
dataset that includes only OTT Subscriptions and OTT Spending.

The data sources are as follows:

\begin{itemize}
    \item \textbf{Società Italiana degli Autori ed Editori (SIAE):}  \\
    Number of screenings in theaters, cinema admissions, ticket prices, audience spending \cite{SIAE}.
    \item \textbf{International Video Federation (IVF) \& Statista:}  \\
    OTT subscriptions and OTT spending (annual data only) \cite{IVF, statista}.
    \item \textbf{Istituto Nazionale di Statistica (ISTAT):}  \\
    Consumer Price Index (CPI) (Month-over-Month) for all items and CPI (MoM) for recreation and culture \cite{CPI}.
    \item \textbf{Istituto Superiore di Sanità (ISS):}  \\
    Number of reported COVID-19 cases \cite{ISS}.
    \item \textbf{Google Trends: }  \\
    Search interest for OTT platforms, used as a proxy for streaming consumption behavior \cite{googtrend}.
\end{itemize}



# Data Exploration \& Preprocessing

## Dataset Overview

This study uses two datasets: a monthly dataset (January 2006 - December
2023) and an annual dataset (2006 - 2023). These datasets provide detailed insights into cinema admissions, audience spending, economic
indicators, and trends related to online streaming services (Netflix,
Amazon Prime Video, etc.).

### Monthly Dataset

The monthly dataset serves as the primary source for time series
modeling. The key variables included are:


\begin{itemize}
    \item \textbf{Cinema Admissions:} The number of individuals attending cinemas each month.
    \item \textbf{Number of Screenings:} The total number of movie screenings conducted per month.
    \item \textbf{Cinema Audience Spending:} The total expenditure by cinema-goers, including ticket sales and other expenses (in euros).
    \item \textbf{Average Spending Per Person:} The average amount spent per visitor in a given month (in euros).
    \item \textbf{Google Trends OTT:} A measure of online search interest in OTT (Over-the-Top) platforms. According to Statista's study on video-on-demand platforms, the most popular SVOD services in Italy as of June 2023 was Netflix, second came Amazon Prime Video, and Disney+ followed behind. \cite{platforms} Therefore Google Trends for these platforms serve as an estimate of consumer engagement with streaming services.
    \item \textbf{CPI (All Items):} The month-over-month Consumer Price Index for all goods and services, reflecting general inflation.
    \item \textbf{CPI (Culture):} The CPI specific to cultural and entertainment-related expenses.
    \item \textbf{COVID-19 Cases:} The number of reported COVID-19 cases per month.
\end{itemize}

### Annual Dataset

The annual dataset provides OTT Subscriptions, OTT Spending in annual
format and aggregated yearly data on cinema attendance. This dataset is
used to examine long-term competition in cinema admissions and OTT
subscriptions. The key variables included in the annual dataset are:

\begin{itemize}
    \item \textbf{OTT Subscriptions:} The total number of active OTT platform subscriptions per year.
    \item \textbf{OTT Spending:} The total consumer expenditure on OTT services (in euros).
    \item \textbf{Cinema Admissions:} The yearly total of cinema admissions.
\end{itemize}



## Handling Missing Values

The dataset initially lacked monthly data for cinema-related features (audience admissions, number of screenings, audience spending, average spending per person) for the year 2013, with only the total annual values available. To address this issue, monthly values were imputed using a proportion-based approach to ensure consistency with historical trends and seasonality. To impute the missing values, monthly averages from the adjacent years (2011, 2012, 2014, and 2015) were used to calculate proportional distributions. These proportions were then applied to the total annual values for 2013.

\begin{figure}[H]
\centering
\includegraphics[width=0.83\textwidth]{plots/missing_vs_imputed-1.png}
\vspace{10pt}
\caption{Original vs. Imputed Monthly Dataset}
\end{figure}

\begin{wrapfigure}[11]{r}{0.60\textwidth}
\vspace{-20pt}
\includegraphics[width=0.58\textwidth]{plots/plot_annual_raw-1.png}
\vspace{-20pt}
\caption{Overview of the Annual OTT Dataset}
\end{wrapfigure}

Additionally, a small number of missing values in 2020 and 2021 (likely due to disrupted operations during the COVID-19 period lock-downs) were replaced with the average ticket price for those respective years.

The annual dataset doesn't have missing values.


## Collinearity Check 

Before proceeding with modeling, it is important to examine potential collinearity issues among the predictor variables. 



### Correlation Analysis for Monthly Data

\begin{wrapfigure}[11]{r}{0.4\textwidth}
\vspace{-40pt}
\includegraphics[width=0.38\textwidth]{plots/plot_corr_monthly-1.png}
\caption{Correlation Matrix for Monthly Data}
\end{wrapfigure}
\vspace{10pt}

To assess the degree of linear relationships between variables, a correlation matrix was computed for the monthly dataset. 


\begin{itemize}
    \item \textbf{Cinema Admissions and Audience Spending} exhibit a very strong positive correlation.
    \item \textbf{Cinema Admissions and Number of Screenings} are also positively correlated.
    \item \textbf{Google Trends OTT and Cinema Admissions} display a moderate negative correlation.
    \item Inflation Indicators (CPI All Items, CPI Culture) show weak correlations with other variables.
\end{itemize}



### Variance Inflation Factor (VIF) Analysis for Monthly Data

\begin{wraptable}[10]{r}{0.35\textwidth}
\centering
\small
\vspace{-20pt}
\begin{tabular}{|l|c|}
    \hline
    \textbf{Variable} & \textbf{VIF Value} \\
    \hline
    Cinema Screenings & 2.13 \\
    Cinema Audience Spending & 1.98 \\
    Cinema Avg Spending Per Person & 1.77 \\
    OTT Google Trends & 1.75 \\
    CPI (All Items) & 1.39 \\
    CPI (Culture) & 1.13 \\
    COVID-19 Cases & 1.58 \\
    \hline
\end{tabular}
\vspace{10pt}
\caption{Variance Inflation Factor (VIF) for Monthly Dataset}
\end{wraptable}
\vspace{10pt}

While the correlation matrix provides valuable insights, it does not explicitly indicate whether collinearity poses a problem for regression models. Therefore, VIF analysis was conducted.


All VIF values are below 5, suggesting that multicollinearity does not pose a major issue for regression modeling. Cinema Screenings has the highest VIF value (2.13), followed by Cinema Audience Spending (1.98), but both remain well within the acceptable range. 


### OTT Google Trends as a Proxy for OTT Subscriptions and Spending

\begin{wraptable}[11]{r}{0.5\textwidth}
\centering
\vspace{-10pt}
    \begin{tabular}{|l|c|}
        \hline
        \textbf{Variable Pair} & \textbf{Correlation Coeff.} \\
        \hline
        OTT Subscriptions vs. OTT Spending & 0.9981 \\
        OTT Subscriptions vs. Google Trends OTT & 0.9379 \\
        OTT Spending vs. Google Trends OTT & 0.9344 \\
        \hline
    \end{tabular}
    \vspace{10pt}
    \caption{Correlation Between Google Trends OTT and OTT Data}
    \label{tab:correlation_ott}
\end{wraptable}

Since OTT Subscriptions and OTT Spending are only available in annual format, this section examines whether Google Trends OTT data can serve as a proxy for monthly modeling. 

The correlation matrix presented in \autoref{tab:correlation_ott} confirms a strong relationship between the variables.

\begin{wrapfigure}[8]{r}{0.48\textwidth}
\vspace{-70pt}
    \includegraphics[width=0.45\textwidth]{plots/compare_ott_trends-1.png}
    \vspace{-10pt}
    \caption{Comparison of OTT Subscriptions, OTT Spending, and Google Trends OTT (Normalized)}
    \label{fig:ott_trends}
\end{wrapfigure}


The results strongly suggest that Google Trends OTT search activity closely follows OTT Subscriptions and Spending trends over time. This supports the assumption that Google Trends data can serve as a reliable proxy for modeling OTT behavior. The visualization suggest that search interest in OTT platforms generally moves in sync with actual market adoption, though there may be a lagged relationship where increases in search activity trigger corresponding rises in subscriptions and spending.

Based on the high correlation values and trend alignment, Google Trends OTT data can be considered a viable proxy for monthly OTT Subscriptions and Spending.


## Time Series Decomposition

Before testing models on the data, an in-depth analysis of the Cinema Admissions time series is performed to examine trend, seasonality, and any possible disruptions.

\begin{wrapfigure}[8]{r}{0.42\textwidth}
\vspace{-50pt}
\centering
\includegraphics[width=0.40\textwidth]{plots/cinema_admissions-1.png}
\caption{Time Series and Autocorrelation Analysis for Cinema Admissions}
\label{fig:cinema_admissions_ts}
\end{wrapfigure}


### Time Series and Autocorrelation Analysis

To have an understanding of the data, the raw time series and its correlograms are plotted in \autoref{fig:cinema_admissions_ts}.

\begin{wrapfigure}[8]{r}{0.4\textwidth}
\vspace{-30pt}
\centering
\includegraphics[width=0.38\textwidth]{plots/seasonplot_admissions-1.png}
\caption{Seasonal Plot: Cinema Admissions (Yearly Comparison)}
\label{fig:cinema_seasonal}
\end{wrapfigure}

\begin{itemize}
    \item The time series exhibits strong seasonality
    \item A downward trend is visible after 2020, indicating a structural break caused by the COVID-19 pandemic.
    \item The correlogram shows significant autocorrelations at seasonal lags, confirming the presence of seasonality.
\end{itemize}


### Seasonal Pattern Analysis

Cinema admissions data reveals distinct seasonal behavior, visualized in Figure \ref{fig:cinema_seasonal}.




\begin{itemize}
    \item Admissions exhibit a recurring seasonal cycle, with peaks in January and December (likely due to holiday season).
    \item Lowest admissions are observed during summer season.
    \item The years 2020 and 2021 deviate significantly from seasonal behavior, indicating the COVID-19 impact.
    \item The recovery in 2022 and 2023 suggests a recovery, but levels remain lower than earlier years.
\end{itemize}

